Fast and adaptive cointegration based model for forecasting high frequency financial time series

### 2016-07-06 articles-conferences/iscef2016/main.tex
agregando imágenes nuevas
change order chapters

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof

\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
\usepackage{amsfonts,amssymb,amsmath,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{multirow,varwidth}
\usepackage{hhline}
\usepackage{url}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{booktabs}
\let\Algorithm\algorithm
\renewcommand\algorithm[1][]{\Algorithm[#1]\setstretch{1.1}}

@@

Cointegration is a long-run property of some non-stationary time series where a 
linear combination of those time series is stationary.
This behaviour has been studied in finance because cointegration restrictions often improve forecasting.
The Vector Error Correction Model (VECM) is a well-known econometric technique that
characterises short-run variations of a set of cointegrated time series incorporating long-run relationships as an error correction term. VECM has been broadly used with low frequency time series. We aimed to adapt VECM to be used in finance with high frequency stream data.
%

Cointegration relations change in time and therefore VECM parameters must be updated when new data is available. We studied how forecasting performance is affected when VECM parameters and the length of historical data used change in time. We observed that the number of cointegration relationships varies with the length of historical data used. Moreover, parameters that increased these relationships in time led to better forecasting performance. Our proposal, called an Adaptive VECM (AVECM) is 
to make a parameters grid search that maximises the number of cointegration relationships in the near past. To ensure the search can be executed fast enough, we used a distributed environment.

%
The methodology was tested using four 10-seconds frequency time series of the Foreign Exchange market. We compared our proposal with ARIMA and the naive forecast of the random walk model. Numerical experiments showed that on average AVECM performed better than ARIMA and random walk. Additionally, our parallel approach AVECM significantly improved execution times with respect to its serial version.

@@

@@

\begin{acknowledgements}
L. Salinas acknowledges support from Basal Project
FB0821 CCTVal-Centro Científico Tecnológico de Valparaíso. P. Arce and W.
Kristjanpoller thank PIIC 2014 DGIP-UTFSM grant for supporting this research.
\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics

\bibliographystyle{plain}       % APS-like style for physics

>>>>

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format) ******
%\usepackage[doublespacing]{setspace}   
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof

%\usepackage{natbib}
\usepackage{graphicx}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
\usepackage{amsfonts,amssymb,amsmath,
            amssymb,amsopn,amsxtra,amstext,amsbsy}
\usepackage{epsfig}
\usepackage[makeroom]{cancel}
\usepackage{array,arydshln}
\usepackage{multirow,varwidth}
\usepackage{hhline}
\usepackage{url}
\usepackage[utf8x]{inputenc}
\usepackage{fancyhdr}
\usepackage{hyperref,wasysym}
\usepackage{color}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{booktabs}
\let\Algorithm\algorithm
\renewcommand\algorithm[1][]{\Algorithm[#1]\setstretch{1.1}}

@@

Cointegration is a long-run property of some non-stationary time series where a
linear combination of those time series is stationary.  This behaviour has been
studied in finance because cointegration restrictions often improve forecasting.
The Vector Error Correction Model (VECM) is a well-known econometric technique
that characterises short-run variations of a set of cointegrated time series
incorporating long-run relationships as an error correction term. VECM has been
broadly used with low frequency time series. We aimed to adapt VECM to be used
in finance with high frequency stream data.
%
Cointegration relations change in time and therefore VECM parameters must be
updated when new data is available. We studied how forecasting performance is
affected when VECM parameters and the length of historical data used change in
time. We observed that the number of cointegration relationships varies with the
length of historical data used. Moreover, parameters that increased these
relationships in time led to better forecasting performance. Our proposal,
called an Adaptive VECM (AVECM) is to make a parameters grid search that
maximises the number of cointegration relationships in the near past. To ensure
the search can be executed fast enough, we used a distributed environment.
%
The methodology was tested using four 10-seconds frequency time series of the
Foreign Exchange market. We compared our proposal with ARIMA and the naive
forecast of the random walk model. Numerical experiments showed that on average
AVECM performed better than ARIMA and random walk. Additionally, AVECM
significantly improved execution times with respect to its serial version.

@@

\begin{acknowledgements}
L. Salinas acknowledges support from Basal Project
FB0821 CCTVal-Centro Científico Tecnológico de Valparaíso. P. Arce and W.
Kristjanpoller thank PIIC 2014 DGIP-UTFSM grant for supporting this research. We also acknowledge to the Fourth International Symposium in Computational Economics and Finance (ISCEF) 2016 for valuable comments and suggestions.
\end{acknowledgements}

@@

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics

%\bibliographystyle{plain}       % APS-like style for physics
